---
title: "Hotel Bookings"
author: "dchia"
date: "9/5/2021"
output: html_document
---
This is a tutorial dataset and code from Julia Silge.
https://juliasilge.com/blog/hotels-recipes/

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, warning = FALSE, 
                      message = FALSE, echo = TRUE, dpi = 180, 
                      fig.width = 8, fig.height = 5)
# load libraries 
library(tidyverse)
library(skimr)
library(GGally)
library(tidymodels)
```

## Exploratory Data Analysis

First, load the data file and this can be obtained from the link above. 

```{r}
# load datafile 
h <- readr::read_csv("hotels_data.csv")
head(h)
```

Use skim function from skimr package, which provides the summary statistics of the variables and can be used on pipeline. 

```{r}
skim(h)
```

The first question to answer is to build a model to predict which hotel stays include children and which do not (out of those bookings that did not cancel).

```{r}
h <- h %>%  
  filter(is_canceled == 0) %>%
  mutate(
    children = case_when(
      children + babies > 0 ~ "children",
      TRUE ~ "none"
    ),
    required_car_parking_spaces = case_when(
      required_car_parking_spaces > 0 ~ "parking",
      TRUE ~ "none"
    )
  ) %>%
  select(-is_canceled, -reservation_status, -babies)

h
```

Hotel stays without children are at least 11x more than those with children. 

```{r}
h %>% 
  count(children)
```
This graph shows the proportion of hotel stays with/without for city hotel vs. resort hotel each month. 

```{r}
h %>%
  mutate(arrival_date_month = factor(arrival_date_month,
    levels = month.name
  )) %>%
  count(hotel, arrival_date_month, children) %>%
  group_by(hotel, children) %>%
  mutate(proportion = n / sum(n)) %>%
  ggplot(aes(arrival_date_month, proportion, fill = children)) +
  geom_col(position = "dodge") +
  scale_y_continuous(labels = scales::percent_format()) +
  facet_wrap(~hotel, nrow = 2) +
  labs(
    x = NULL,
    y = "Proportion of hotel stays",
    fill = NULL
  )
```
This bar chart shows the proportion of hotel stays with/without children that required car parking spaces. 

```{r}
h %>%  
  count(hotel, required_car_parking_spaces, children) %>%  
  group_by(hotel, children) %>% 
  mutate(proportion = n / sum(n)) %>% 
  ggplot(aes(required_car_parking_spaces, proportion, fill = children)) + 
  geom_col(position = "dodge") + 
  scale_y_continuous(labels = scales::percent_format()) + 
  facet_wrap(~hotel, nrow = 2) + 
  labs(
    x = NULL, 
    y = "Proportion of hotel stays",
    fill = NULL
  )
```
Use ggpairs function to get a high-level view of how variables are related to each other. 

```{r}
h %>% 
  select(
    children, adr, 
    required_car_parking_spaces,
    total_of_special_requests
  ) %>% 
  ggpairs(mapping = aes(color = children))
```
Next, let's create dataset for modelling and convert those columns (characters) we are interested, to factors. 

```{r}
set.seed(1234)

h_df <- h %>% 
  select(
    children, hotel, arrival_date_month,meal, adr, adults,
    required_car_parking_spaces, total_of_special_requests,
    stays_in_week_nights, stays_in_weekend_nights
  ) %>% 
  mutate_if(is.character, factor)

h_df
```

Then build a set of steps for data processing and feature engineering. 
Recipe = what model, and training data
Down sample the data, because there are 11x more hotel stays w/o children
Convert factor columns into binary 0, 1
Remove any numeric variables that have zero variance
Normalize the numeric variables 

Recipe and Prep = do something with the steps and training data, estimate the required parameters from train to implement these steps so the sequence can be applied later to another dataset.
Recipe = contains all transformations for data processing and feature engineering, as well as these transformations were estimated from. 

```{r}
set.seed(1234)

h_split <- initial_split(h_df) 
h_train <- training(h_split)
h_test <- testing(h_split)

h_rec <-  recipe(children ~ ., data = h_train) %>%
  step_downsample(children) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>%
  step_zv(all_numeric()) %>%
  step_normalize(all_numeric()) %>%
  prep()

h_rec
```

```{r}
test_proc <- bake(h_rec, new_data = h_test)
test_proc
```

### KNN 
Then move on to specify and fit our models. 

Juice() = squeeze the training data back out, transformed in the ways we specificied including the downsampling. 

```{r}
knn_spec <- nearest_neighbor() %>%
  set_engine("kknn") %>%
  set_mode("classification")

knn_fit <- knn_spec %>%
  fit(children ~ ., data = juice(h_rec))

knn_fit
```

### Decision Tree 

```{r}
tree_spec <- decision_tree() %>% 
  set_engine("rpart") %>% 
  set_mode("classification")

tree_fit <- tree_spec %>%  
  fit(children ~., data = juice(h_rec))

tree_fit
```

# Evaluate models 

To evaluate the models, we need to build a validation set. We can build a set of Monte Carlo splits from the downsampled training data, and use this set of resamples to estimate the performance of the two models. 

```{r}
set.seed(1234)
validation_splits <- mc_cv(juice(h_rec), prop = 0.9, strata = children)
validation_splits
```


```{r}
knn_res <- fit_resamples(
  knn_spec,
  children ~ .,
  validation_splits,
  control = control_resamples(save_pred = TRUE)
)

knn_res %>%
  collect_metrics()
```
```{r}
tree_res <- fit_resamples(
  tree_spec,
  children ~ .,
  validation_splits,
  control = control_resamples(save_pred = TRUE)
)

tree_res %>% 
  collect_metrics()
```


```{r}
knn_res %>% 
  unnest(.predictions) %>% 
  mutate(model = "kknn") %>% 
  bind_rows(tree_res %>% 
              unnest(.predictions) %>% 
              mutate(model = 'rpart')) %>% 
  group_by(model) %>% 
  roc_curve(children, .pred_children) %>% 
  ggplot(aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line(size = 1.5) +
  geom_abline(
    lty = 2, alpha = 0.5, 
    color = "gray50",
    size = 1.2
  )
```

```{r}
knn_conf <- knn_res %>% 
  unnest(.predictions) %>% 
  conf_mat(children, .pred_class)

knn_conf
```

```{r}
knn_conf %>% 
  autoplot()
```

```{r}
knn_fit %>% 
  predict(new_data = test_proc, type = "prob") %>% 
  mutate(truth = h_test$children) %>% 
  roc_auc(truth, .pred_children)
```


















